---
title: "Overview of the etc5523nhu Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Overview of the etc5523nhu Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

The **etc5523nhu** package provides data and an interactive Shiny app for exploring the burden of healthcare-associated infections (HAIs) in Germany and the EU/EEA, based on *Zacher et al. (2019)*.

It was developed as part of **ETC5523 – Communicating with Data: Assignment 4**.

------------------------------------------------------------------------

# Data

The dataset `hai_burden` was created from *Table 2 in Zacher et al. (2019)* using R.\
It was first defined manually with `tibble::tribble()` to include estimated values for each infection type and measure, then reshaped to long format using `tidyr::pivot_longer()`.

I also added simulated uncertainty intervals (`ci_low` and `ci_high`) to make the plots more realistic and visually informative.

Finally, the dataset was saved to the package with `usethis::use_data(hai_burden, overwrite = TRUE)`.

```{r}
library(etc5523nhu)
data("hai_burden")
knitr::kable(head(hai_burden), caption = "Preview of the hai_burden dataset")

```

# The Shiny App

Launch the interactive dashboard with:

```{r, eval=FALSE}
run_app()
```

The app allows users to:

-   Select infection types and compare metrics

-   Explore differences between German and EU/EEA samples

-   Interpret results through dynamic plots and explanatory text

# Reference

Zacher B., Haller S., Willrich N., Walter J., Abu Sin M., Cassini A., Plachouras D., Suetens C., & Eckmanns T. (2019). Application of a new methodology and R package reveals a high burden of healthcare-associated infections in Germany compared to the average in the EU/EEA, 2011–2012. EuroSurveillance, 24(46), 1900135. <https://www.eurosurveillance.org/content/10.2807/1560-7917.ES.2019.24.46.1900135>
